Clearance Project - 3 - TKstats

### 2025-03-24 data-raw/ClearanceProject-3-TKstats.Rmd
Added figure investigating interspecies trends in fup and CLint.

interspecies_rmsle_trends <- bind_rows(rpr_rmsle, rph_rmsle) %>%
  mutate(
    close_clint = ifelse(Chemical %in% close_clint,
                         "|\U0394 CLint| within 50", "|\U0394 CLint| > 50")
  ) %>% 
  filter(Chemical %in% unique(rpr_rmsle$Chemical)) %>% 
  pivot_longer(cols = contains("_rmsle"),
               names_to = "clearance_type",
               values_to = "RMSLE") %>%
  pivot_wider(names_from = fupSpecies,
              values_from = RMSLE) %>%
  mutate(norm_diff = (human - rat)/rat) %>%
  filter(!is.na(norm_diff)) %>% 
  group_by(Chemical, close_clint, clearance_type) %>%
  ggplot(aes(x = forcats::fct_reorder(Chemical, norm_diff, mean),
             y = norm_diff,
             fill = clearance_type)) +
  geom_col(
    width = 0.6,
    # fill = "grey5", 
    position = position_dodge(),
    color = NA
    ) +
  labs(
    x = "Chemical",
    y = "Fold Change RMSLE\nHuman more accurate <<     >> Rat more accurate"
  ) +
  scale_fill_manual(values = c("grey20", "seagreen"),
                    breaks = c("nonrest_rmsle", "rest_rmsle"),
                    labels = c("Nonrestrictive", "Restrictive"),
                    name = "Clearance") +
  geom_hline(yintercept = 0) +
  theme_minimal() + 
  facet_grid(cols = vars(close_clint),
             # rows = vars(clearance_type),
             space = "free_x", scales = "free_x") +
  theme(axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(color = "grey5"),
        panel.background = element_rect(color = "grey5"),
        axis.title = element_text(size = 10),
        legend.position = "bottom")

species_clint_pl + interspecies_rmsle_trends +
  plot_layout(widths = c(5,8))

@@

>>>>

interspecies_rmsle_trends <- bind_rows(rpr_rmsle, rph_rmsle) %>%
  mutate(
    close_clint = ifelse(Chemical %in% close_clint,
                         "|\U0394 CLint| within 50", "|\U0394 CLint| > 50")
  ) %>% 
  filter(Chemical %in% unique(rpr_rmsle$Chemical)) %>% 
  pivot_longer(cols = contains("_rmsle"),
               names_to = "clearance_type",
               values_to = "RMSLE") %>%
  pivot_wider(names_from = fupSpecies,
              values_from = RMSLE) %>%
  mutate(norm_diff = (human - rat)/rat)

irt_pl <- interspecies_rmsle_trends %>%
  filter(!is.na(norm_diff)) %>% 
  ggplot(aes(x = forcats::fct_reorder(Chemical, norm_diff, mean),
             y = norm_diff,
             fill = clearance_type)) +
  geom_col(
    width = 0.6,
    # fill = "grey5", 
    position = position_dodge(),
    color = NA
    ) +
  labs(
    x = "Chemical",
    y = "Fold Change RMSLE\nHuman more accurate <<     >> Rat more accurate"
  ) +
  scale_fill_manual(values = c("grey20", "seagreen"),
                    breaks = c("nonrest_rmsle", "rest_rmsle"),
                    labels = c("Nonrestrictive", "Restrictive"),
                    name = "Clearance") +
  geom_hline(yintercept = 0) +
  theme_minimal() + 
  facet_grid(cols = vars(close_clint),
             # rows = vars(clearance_type),
             space = "free_x", scales = "free_x") +
  theme(axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(color = "grey5"),
        panel.background = element_rect(color = "grey5"),
        axis.title = element_text(size = 10),
        legend.position = "bottom")

species_clint_pl + irt_pl +
  plot_layout(widths = c(5,8))

@@

From this we notice that CLint specifically can have drastic differences which
we want to explore. We also glean that there are potential drastic improvements
in model accuracy. We next want to compare how a changing human CLint towards rat
CLint by 1% steps changes RMSE.

```{r}
combo_clint <- rpr_param_df %>%
  select(Chemical, rCLint = CLint) %>% 
  left_join(rph_param_df %>%
              select(Chemical, hCLint = CLint)) %>%
  mutate(hclint_1p = (1/10)*hCLint, 
         direction = sign(hCLint - rCLint)) %>%
  mutate(Nstep = ifelse(direction == 0, 
                        0, abs(hCLint - rCLint) %/% hclint_1p))

combo_clint %>%
  left_join(interspecies_rmsle_trends) %>%
  ggplot(aes(Nstep, norm_diff)) +
  geom_point() +
  scale_x_log10() +
  theme_bw()
```
